---
title: "Stats140XP_Final Project"
author: "Group 2"
date: "11/22/2022"
output: pdf_document
---

# Deciding what we want to look at

We want to find out good risk factors of heart disease. Since we are not experts in medical terms, we 

# EDA

```{r}
heart <- read.csv(file = 'heart.csv', stringsAsFactors = T)
library(ggcorrplot)
heart$Sex <- as.numeric(heart$Sex)
heart$HeartDisease <- as.numeric(heart$HeartDisease)
index_to_investigate <- c("Age", "Sex", "RestingBP", "FastingBS", "MaxHR", "HeartDisease", "Cholesterol")
heart.investigated<- heart[index_to_investigate]
head(heart.investigated)

```

```{r}
corr <- cor(heart[,index_to_investigate])
p.mat <- cor_pmat(heart[,index_to_investigate])
ggcorrplot(corr,
           hc.order = TRUE,
           type = "lower",
           p.mat = p.mat)
```

From the correlation matrix, we mainly focus on corelations between variables and Heart Disease. we can see that max Heart Rate and Cholesterol have a  negative correlation with Heart disease, while sex, age, Fasting Blood Sugar level, and resting blood pressure has positive correlation with Heart Disease.

We can also look at the correlation between the interaction terms. However, they are not the main concern of this study.

We then look at the distribution for each variable to observe patterns and abnormalties.

# Visualizing HeartDisease and a continuous variable.
```{r}
heart <- read.csv(file = 'heart.csv', stringsAsFactors = T)
heart$HeartDisease <- as.character(heart$HeartDisease)

# Between the ages of 55 and 65 is the high incidence period of Heart Failure.
ggplot(data = heart, mapping = aes(x = Age)) + geom_freqpoly(mapping = aes(colour = HeartDisease), binwidth = 1)
above_50 <- heart$Age >= 50
cat("People above 60 is ", (mean(as.integer(heart$HeartDisease[above_50]))/mean(as.integer(heart$HeartDisease[!above_50]))-1)*100, "% more likely to get heart disease", sep = "")


# RestingBP is not a significant factor to influence Heart Failure.
ggplot(data = heart, mapping = aes(x = RestingBP)) + geom_freqpoly(mapping = aes(colour = HeartDisease), binwidth = 10)
BP_above_145 <- heart$RestingBP >= 145
cat("People with Resting Blood pressure above 145 is ", (mean(as.integer(heart$HeartDisease[BP_above_145]))/mean(as.integer(heart$HeartDisease[!BP_above_145]))-1)*100, "% more likely to get heart disease", sep = "")

# Cholesterol
ggplot(data = heart, mapping = aes(x = Cholesterol)) + geom_freqpoly(mapping = aes(colour = HeartDisease), binwidth = 10)
```

Note that there are many 0s recorded for Cholesterol level, which should indicate Non-recorded entries, i.e. NAs. Thus we remove the 0s, and new distribution look like this:

```{r}
NA.index <- heart$Cholesterol == 0
heart$Cholesterol[NA.index] <- NA
ggplot(data = heart, mapping = aes(x = Cholesterol)) + geom_freqpoly(mapping = aes(colour = HeartDisease), binwidth = 10)

cho_above_280 <- heart$Cholesterol >= 280
cat("People with Cholesterol above 280 are ", (mean(as.integer(heart$HeartDisease[cho_above_280]), na.rm = T)/mean(as.integer(heart$HeartDisease[!cho_above_280]), na.rm = T)-1)*100, "% more likely to get heart disease", sep = "")
```


```{r}
# Higher MaxHR is associated with higher risks of Heart Failure.
ggplot(data = heart, mapping = aes(x = MaxHR)) + geom_freqpoly(mapping = aes(colour = HeartDisease), binwidth = 10)

MaxHR_below_150 <- heart$MaxHR <= 150
cat("People with Max Heart Rate below 150 are ", (mean(as.integer(heart$HeartDisease[MaxHR_below_150]), na.rm = T)/mean(as.integer(heart$HeartDisease[!MaxHR_below_150]), na.rm = T)-1)*100, "% more likely to get heart disease", sep = "")
```

# Visualizing HeartDisease and a categorical variable.
```{r}
library(ggplot2)
library(tidyverse)
# The incidence of heart failure is higher in women.
ggplot(data = heart) +
  geom_count(mapping = aes(x = Sex, y = HeartDisease, colour = Sex))

# The incidence of heart failure is higher in FastingBS > 120 mg/dl.
heart$FastingBS <- ifelse(heart$FastingBS == 1, ">120 mg/dl", "<120 mg/dl")
ggplot(data = heart) +
  geom_count(mapping = aes(x = FastingBS, y = HeartDisease, colour = FastingBS))

FastingBS_above_120 <- heart$FastingBS == ">120 mg/dl"
cat("People with Max Heart Rate below 150 are ", (mean(as.integer(heart$HeartDisease[FastingBS_above_120]), na.rm = T)/mean(as.integer(heart$HeartDisease[!FastingBS_above_120]), na.rm = T)-1)*100, "% more likely to get heart disease", sep = "")
```


# ANOVA

use ANOVA to check whether patterns we observed are are significant.

```{r}
anova1 <- aov(HeartDisease~.,heart[,index_to_investigate])
summary(anova1)
anova2 <- aov(HeartDisease~FastingBS,heart[,index_to_investigate])
summary(anova2)
anova2 <- aov(HeartDisease~Sex,heart[,index_to_investigate])
summary(anova2)
anova2 <- aov(HeartDisease~Age,heart[,index_to_investigate])
summary(anova2)
anova2 <- aov(HeartDisease~Cholesterol,heart[,index_to_investigate])
summary(anova2)
anova2 <- aov(HeartDisease~MaxHR,heart[,index_to_investigate])
summary(anova2)
anova2 <- aov(HeartDisease~RestingBP,heart[,index_to_investigate])
summary(anova2)
```

We conclude that the only insignificant indicator is Fasting Blood Sugar Level.


# Bonus: Graphs used on poster

```{r}
heart <- read.csv(file = 'heart.csv', stringsAsFactors = T)
library(tidyverse)
#install.packages("ggalluvial")
library(ggalluvial)

 heart %>% group_by(RestingECG,ST_Slope,HeartDisease) %>% tally() %>% mutate(prop = n/sum(n)) %>% ggplot(.)+
    geom_col(aes(RestingECG, prop, fill = HeartDisease)) +
    facet_wrap(~ST_Slope)

heart %>% group_by(RestingECG,ST_Slope,HeartDisease) %>% tally() %>% mutate(prop = n/sum(n)) %>% ggplot(.)+
    geom_col(aes(RestingECG, prop, fill = HeartDisease)) +
    facet_wrap(~ST_Slope)

ggplot(data = heart,
       aes(axis1 = Sex, axis2 = ChestPainType, axis3 = RestingECG,axis4 = ExerciseAngina,axis5=ST_Slope,
           y = MaxHR)) +
  scale_x_discrete(limits = c("Sex", "ChestPainType", "RestingECG","ExerciseAngina","ST_Slope")) +
  xlab("categorical varibales") +
  geom_alluvium(aes(fill = HeartDisease)) +
  geom_stratum() +
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  theme_minimal() +
  ggtitle("") + coord_flip()
```

